### 2024-11-24 paper/paper.qmd
draft of title, abstract, introduction
draft of title, abstract, introduction

---
title: "Analysis and Prediction of Shared Bicycle Usage at the University of Toronto St. George Campus"
subtitle: "Main Finding TBD"
author: 
  - Haowei Fan
thanks: "Code and data are available at: https://github.com/HaoweiFan0912/Bikeshare-Forecast.git"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

#### Read data ####

start <- read_parquet(here::here("data/02-analysis_data/start.parquet"))
stop <- read_parquet(here::here("data/02-analysis_data/stop.parquet"))
```

```{r}

```

# Introduction

Overview paragraph

Estimand paragraph

Results paragraph

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....

>>>>

---
title: "Analysis and Prediction of Shared Bicycle Usage at the University of Toronto St. George Campus"
subtitle: "Main Finding TBD"
author: 
  - Haowei Fan
thanks: "Code and data are available at: https://github.com/HaoweiFan0912/Bikeshare-Forecast.git"
date: today
date-format: long
abstract: "Students and staff at the University of Toronto's St. George campus often face difficulties finding a bike-sharing parking spot when arriving at campus, and finding a bike when leaving, which highlights the need for effective prediction of future bike-sharing demand to support campus commuting infrastructure. This study uses Bayesian Poisson regression to predict the usage of bike-sharing stations at 27 locations on campus for specific years, months, and  four-hour intervals of the day. The results show that bike-sharing usage on campus has been increasing annually, with significant seasonal variations and peak usage between 8 am and 8 pm."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

#### Read data ####
start <- read_parquet(here::here("data/02-analysis_data/start.parquet"))
stop <- read_parquet(here::here("data/02-analysis_data/stop.parquet"))
```

# Introduction
Bike-sharing services have undoubtedly improved commuting efficiency, especially for short trips within a campus. For example, a study on bike-sharing in Xiamen, China, highlighted how data-driven analyses can inform policymakers about system performance, thereby guiding investments and policy decisions to enhance urban mobility[https://arxiv.org/abs/2401.03987?utm_source=chatgpt.com]. However, in the Toronto region, despite a study by the University of Toronto's School of Cities showing that ridership increased from about 665,000 trips in 2015 to over 4.5 million in 2022[https://schoolofcities.github.io/bike-share-toronto/growth], there has been little research focused on small-scale areas, such as universities, within Toronto. Furthermore, a study from Hangzhou, China, confirmed the significant differences in traveling characteristics between cities and university campuses[https://file.techscience.com/uploads/attached/file/20200916/20200916072445_75552.pdf?utm_source=chatgpt.com]. Therefore, this paper takes the University of Toronto as a case study to analyze bike-sharing usage within the campus.

本研究的核心在于，从自2017年月1日到2024年9月30日中的所有共享单车使用数据中抽取多伦多大学校园内27个站台的样本，并按照4小时的区间，计算每个站台在各个区间内的共享单车使用总量。并且通过贝叶斯柏松回归模型根据站台，年份，月份，日期，以及这一天的具体4小时区间，来预测该站台在此区间内的使用数量。

本研究发现，尽管多大校园内所有站点的使用量每年都呈现极速上涨，但都存在巨大的季节性差异。冬天的使用量无论年份，站点统统呈现着接近0的使用量。另外，在每年的1-9月，使用量呈上升趋势，之后的几个月逐渐下降。一天中不同时间段的使用量也存在明显差异其中8am到8pm的使用量平均占了一天中的79.8%，相比之下0am-8am的使用量只有6.8%。另外值得注意的是27个站台中单车的开出量于停放量基本持平，但存在4个站台他们的进出量纯在较大差异。

这些见解相比于笼统的大多伦多地区共享单车使用数量的统计，更能针对多伦多大学校区内的交通建设以及政策调整，以及如何调度，投放或储存共享单车提供更详细的建议。

结构：TBD

## Estimand
本研究旨在估计特定4小时区间内多伦多大学校园内某个共享单车站台的使用量。通过将时间转换为年，月，日，这一天特定的4小时区间，来考虑到某个站台使用量的总体趋势，季节以及hourly variation。核心目标是探索多伦多大学校园内27个共享单车站台使用量随时间的变化，从而给政策制定者建议以调度自行车，来提高校园内的通勤效率。

### 2024-11-24 scripts/05-exploratory_data_analysis.qmd
draft of title, abstract, introduction
draft of title, abstract, introduction

```{r}
#| warning: false
#| message: false
start_asdate <- start 
start_asdate$time <- as.Date(start_asdate$time)

daily_counts <- start_asdate %>%
  group_by(time) %>%
  summarise(total_count = sum(count))

ggplot(data = daily_counts, aes(x = time, y = total_count)) +
  geom_point() +
  labs(title = "Daily usage from January 2017 to September 2024", x = "Time", y = "Count") +
  theme_minimal()+  
  geom_smooth(method = "lm")  
```

@@

total_counts <- start_astime %>%
  group_by(time_interval) %>%
  summarise(count = sum(count, na.rm = TRUE))

ggplot(data = total_counts, aes(x = time_interval, y = count, fill = time_interval)) +
  geom_bar(stat = "identity") +
  labs(title = "Total usage during different time periods from January 2017 to September 2024", 
       x = "Time interval", y = "Count") +
  theme_minimal() +
  theme(legend.position = "none") 
```
```{r, fig.width=10}
options(scipen = 999)

@@

ggplot(merged_data, aes(x = station_name, y = count_difference)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title = "Difference of daily average of bicycle parking and departure volumes for each station",
       x = "Station Name",
       y = "Parking volume minus departure volume")+
  ylim(-4, 3)+
  coord_flip()
```

>>>>

```{r}
#| warning: false
#| message: false
start_asdate <- start 
start_asdate$time <- as.Date(start_asdate$time)
daily_counts <- start_asdate %>%
  group_by(time) %>%
  summarise(total_count = sum(count))
ggplot(data = daily_counts, aes(x = time, y = total_count)) +
  geom_point() +
  labs(title = "Daily usage from January 2017 to September 2024", x = "Time", y = "Count") +
  theme_minimal()+  
  geom_smooth(method = "lm")  
```

@@

total_counts <- start_astime %>%
  group_by(time_interval) %>%
  summarise(count = sum(count, na.rm = TRUE)) %>%
  mutate(percentage = count / sum(count) * 100)

# 绘制条形图并添加百分比标签
ggplot(data = total_counts, aes(x = time_interval, y = count, fill = time_interval)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = paste0(round(percentage, 1), "%")), 
            vjust = -0.5, size = 3) + # 添加百分比标签
  labs(title = "Total usage during different time periods from January 2017 to September 2024", 
       x = "Time interval", y = "Count") +
  theme_minimal() +
  theme(legend.position = "none")
```
```{r, fig.width=10}
options(scipen = 999)

@@

ggplot(merged_data, aes(x = station_name, y = count_difference)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title = "Difference of daily average of bicycle parking and departure volumes for each station",
       x = "Station Name",
       y = "Parking volume minus departure volume")+
  ylim(-4, 3)+
  coord_flip()
```
```{r}
df_month <- start
df_month$time <- as.POSIXct(df_month$time, format = "%Y-%m-%d %H:%M:%S")
df_month <- df_month %>% select(-station_name)
# 提取年月信息并添加新列
df_month$year_month <- format(df_month$time, "%m")

# 按照年月和站点分组并汇总 count
monthly_count <- df_month %>%
  group_by(year_month) %>%
  summarise(total_count = sum(count)/30)

# 绘制每月所有站点的总和柱状图
ggplot(monthly_count, aes(x = year_month, y = total_count)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(title = "每个月所有站台的 Count 总和",
       x = "年月",
       y = "总和 Count",
       fill = "站台名称") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
